---
title: "IS 607 Assignment 5"
author: "Derek G. Nokes"
date: "Friday, February 27, 2015"
output: pdf_document
toc: yes
---

\pagebreak

## Connect to PostgreSQL
We connect to the PostgreSQL database as follows:

```{r}
# load the RPostgreSQL library
library('RPostgreSQL')

# connect to PoststgreSQL using RPostgreSQL
driverHandle <- dbDriver("PostgreSQL")
 
# establish connection to 'flights' database
connectionHandle <- dbConnect(driverHandle, 
                              dbname="flights",
                              host="localhost",
                              port=5432,user="postgres",
                              password  ="tgdnrx78")
```

## Query PostgreSQL

We extract the flights data from any of the three major NYC airports to 'Los Angeles Intl' airport during the period between 2013-01-23 and 2013-03-01 and includes the 
following information:

* Date of Departure: d_date
* Hour of departure: hour
* Airport of origin: o_airport
* Airport of destination: d_airport
* Carrier: carrier
* Approximate temperature at scheduled time of departure: d_temp
* Departure delay (in minutes): d_delay
* Arrival delay (in minutes): a_delay
* Air time (in minutes): air_time
* Seating capacity of the airplane: seating 

```{r}
# build the query (join the flights,planes,airports,and weather tables)
query<-"SELECT (f.year::text || '-' || f.month::text || '-' || f.day::text)::date 
as d_date,w.hour,o.name as o_airport,d.name as d_airport,
f.carrier,w.temp as d_temp,
f.dep_delay as d_delay,arr_delay as a_delay,air_time,
p.seats as seating FROM flights f
INNER JOIN (SELECT * FROM planes) p ON f.tailnum = p.tailnum
INNER JOIN (SELECT * FROM airports) d ON f.dest =d.faa
INNER JOIN (SELECT * FROM airports) o ON f.origin = o.faa
INNER JOIN (SELECT concat_ws('-',weather.year,weather.month,
weather.day,weather.hour) as w_date_hour,weather.* FROM weather) w 
ON concat_ws('-',f.year,f.month,f.day,f.hour) = w.w_date_hour AND 
f.origin = w.origin
WHERE (d.name='Los Angeles Intl')
AND ((f.year::text || '-' || f.month::text || '-' || f.day::text)::date 
BETWEEN '2013-02-23' AND '2013-03-01')"
# execute the query
data<-dbGetQuery(connectionHandle,query)
```

## Data Frame Object Properties

We can look at the properties of the data frame as follows:

```{r}
# look at the characteristics of the data data.frame object
str(data)
# look at the first 10 rows of the data frame
head(data,10)

```

## Data Frame Contents
The following data table contains the data:

```{r}
# creat the table
knitr::kable(data, caption = 'NYC flights to Los Angeles Intl Between 2013-02-23 and 2013-03-01')
```

## Disconnect From PostgreSQL

Disconnect from the database:

```{r}
# Close PostgreSQL connection 
dbDisconnect(connectionHandle)
```


